Word-Referent Identification Under Multimodal Uncertainty

### 2018-02-24 Writeup/ms.tex
second draft journal write-up

---
title             : "Word-Referent Identification Under Multimodal Uncertainty"
shorttitle        : "Word Identification Under Multimodal Uncertainty"

author: 
  - name          : "Abdellah Fourtassi"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "afourtas@stanford.edu"
  - name          : "Michael C. Frank"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Department of Psychology, Stanford University"

author_note: |

  Abdellah Fourtassi

@@

abstract: |

 Identifying a spoken word in a referential context requires both the ability to process and integrate multimodal input and the ability to reason under uncertainty. How do these tasks interact with one another? We introduce a task that allows us to examine how adults identify words under joint uncertainty in the auditory and visual modalities. We propose an ideal observer model of the task which provides an account of how auditory and visual cues are combined optimally. Model predictions are tested in three experiments where word recognition is made under two kinds of uncertainty: category ambiguity and/or distorting noise. In all cases, the optimal model explains much of the variance in human mean judgments. In particular, when the signal is not distorted with noise, participants weight the auditory and visual cues optimally, that is, according to the relative reliability of each modality. But when one modality has noise added to it, human perceivers systematically prefer the unperturbed modality to a greater extent than the optimal model does. The study provides a formal framework which helps us understand precisely how word form and word meaning interact in word recognition under uncertainty. Moreover it offers a first step towards a model that accounts for form-meaning synergy in early word learning.

keywords          : "Language understanding; audio-visual processing; word learning; speech perception; computational modeling."

wordcount         : "X"

header-includes:
   #- \usepackage{bibentry}
   - \usepackage[sortcites=false,sorting=none]{biblatex}

bibliography      : ["references.bib"]

figsintext        : no
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes

lang              : "english"
class             : "man"
output            : papaja::apa6_pdf

citation_package: biblatex

---

```{r load_packages, include = FALSE}
library(papaja)
library(readr)
library(tidyr)
library(ggplot2)
library(cowplot)
library(knitr)
library(boot)
library(dplyr)
library(langcog)
theme_set(theme_bw())
library(broom)
#library("knitcitations")
#cleanbib()
#options("citation_format" = "pandoc")
```

```{r}

#Data from the 3 experiments 
#############################
exp1 <- read_delim("../Data_and_analysis/data_exp1_anonym.txt", delim = " ") %>%
  filter(type == "Task") %>%
  mutate(experiment='Experiment1')

exp2 <- read_delim("../Data_and_analysis/data_exp2_anonym.txt", delim = " ") %>%
  filter(type == "Task") %>%
  mutate(experiment='Experiment2')

exp3 <- read_delim("../Data_and_analysis/data_exp3_anonym.txt", delim = " ") %>%
  filter(type == "Task") %>%
  mutate(experiment='Experiment3')
```

```{r}
summarise <- dplyr::summarise
rename <- dplyr::rename
```

```{r}
#First Exlusion criteria:
#############################

#All data
N_all_1 <- exp1 %>%
  distinct(ID) %>%
  nrow()

N_all_2 <- exp2 %>%
  distinct(ID) %>%
  nrow()

N_all_3 <- exp3 %>%
  distinct(ID) %>%
  nrow()

#Participants who did not encounter a technical problem with the online experiment

#Exp1
exp1_noProb <- exp1 %>%
  filter(problem=="No")

N_noProb_1 <- exp1_noProb %>%
  distinct(ID) %>%
  nrow()

#Exp2
exp2_noProb <- exp2 %>%
  filter(problem=="No")

N_noProb_2 <- exp2_noProb %>%
  distinct(ID) %>%
  nrow()

#Exp3
exp3_noProb <- exp3 %>%
  filter(problem=="No")

N_noProb_3 <- exp3_noProb %>%
  distinct(ID) %>%
  nrow()

#Participants who who did not encounter a problem AND were above 50% accuracy on obvious trials

#Exp1
exp1_good <- exp1_noProb %>%
  filter(score > 0.5)

N_good_1 <- exp1_good %>%
  distinct(ID) %>%
  nrow()

#Exp2
exp2_good <- exp2_noProb %>%
  filter(score > 0.5)

N_good_2 <- exp2_good %>%
  distinct(ID) %>%
  nrow()

#Exp3
exp3_good <- exp3_noProb %>%
  filter(score > 0.5)

N_good_3 <- exp3_good %>%
  distinct(ID) %>%
  nrow()

```

```{r}

#Data subseting

#Experiment 1
##############

sound_all_exp1 <- exp1_good %>%
    filter(condition == "sound")

concept_all_exp1 <- exp1_good %>%
    filter(condition == "concept")

joint_all_exp1 <- exp1_good %>%
    filter(condition == "joint")

#Summary
sounds_exp1 <- exp1_good %>%
  filter(condition == "sound") %>%
  group_by(sound_dist) %>%
  dplyr::summarise(mean = mean(answer),
                   sd = sd(answer),
                   n = n()) %>%
  mutate(se = sd / sqrt(n),
         lower = mean - qt(1 - (0.05 / 2), n - 1) * se,
         upper = mean + qt(1 - (0.05 / 2), n - 1) * se) %>%
  select(-sd, -n, -se,) %>%
  mutate(Experiment="Experiment 1", 
         Condition="Auditory") %>%
  rename(distance = sound_dist)

concepts_exp1 <- exp1_good %>%
  filter(condition == "concept") %>%
  group_by(concept_dist) %>%
  dplyr::summarise(mean = mean(answer),
                   sd = sd(answer),
                   n = n()) %>%
  mutate(se = sd / sqrt(n),
         lower = mean - qt(1 - (0.05 / 2), n - 1) * se,
         upper = mean + qt(1 - (0.05 / 2), n - 1) * se) %>%
  select(-sd, -n, -se,) %>%
  mutate(Experiment="Experiment 1", 
         Condition="Visual") %>%
  rename(distance = concept_dist)

joint_exp1 <- exp1_good %>%
  filter(condition == "joint") %>%
  group_by(concept_dist, sound_dist) %>%
  summarise(mean = mean(answer)) %>%
  mutate(Experiment="Experiment 1", 
         Condition="Joint")

#Experiment2
############

sound_all_exp2 <- exp2_good %>%
    filter(condition == "sound")

concept_all_exp2 <- exp2_good %>%
    filter(condition == "concept")

joint_all_exp2 <- exp2_good %>%
    filter(condition == "joint")

#Summary
sounds_exp2 <- exp2_good %>%
  filter(condition == "sound") %>%
  group_by(sound_dist) %>%
  dplyr::summarise(mean = mean(answer),
                   sd = sd(answer),
                   n = n()) %>%
  mutate(se = sd / sqrt(n),
         lower = mean - qt(1 - (0.05 / 2), n - 1) * se,
         upper = mean + qt(1 - (0.05 / 2), n - 1) * se) %>%
  select(-sd, -n, -se,) %>%
  mutate(Experiment="Experiment 2", 
         Condition="Auditory") %>%
  rename(distance = sound_dist)

concepts_exp2 <- exp2_good %>%
  filter(condition == "concept") %>%
  group_by(concept_dist) %>%
  dplyr::summarise(mean = mean(answer),
                   sd = sd(answer),
                   n = n()) %>%
  mutate(se = sd / sqrt(n),
         lower = mean - qt(1 - (0.05 / 2), n - 1) * se,
         upper = mean + qt(1 - (0.05 / 2), n - 1) * se) %>%
  select(-sd, -n, -se,) %>%
  mutate(Experiment="Experiment 2", 
         Condition="Visual") %>%
  rename(distance = concept_dist)

joint_exp2 <- exp2_good %>%
  filter(condition == "joint") %>%
  group_by(concept_dist, sound_dist) %>%
  summarise(mean = mean(answer)) %>%
  mutate(Experiment="Experiment 2", 
         Condition="Joint")

#Experiment3
############

sound_all_exp3 <- exp3_good %>%
    filter(condition == "sound")

concept_all_exp3 <- exp3_good %>%
    filter(condition == "concept")

joint_all_exp3 <- exp3_good %>%
    filter(condition == "joint")

#Summary
sounds_exp3 <- exp3_good %>%
  filter(condition == "sound") %>%
  group_by(sound_dist) %>%
  dplyr::summarise(mean = mean(answer),
                   sd = sd(answer),
                   n = n()) %>%
  mutate(se = sd / sqrt(n),
         lower = mean - qt(1 - (0.05 / 2), n - 1) * se,
         upper = mean + qt(1 - (0.05 / 2), n - 1) * se) %>%
  select(-sd, -n, -se,) %>%
  mutate(Experiment="Experiment 3", 
         Condition="Auditory") %>%
  rename(distance = sound_dist)

concepts_exp3 <- exp3_good %>%
  filter(condition == "concept") %>%
  group_by(concept_dist) %>%
  dplyr::summarise(mean = mean(answer),
                   sd = sd(answer),
                   n = n()) %>%
  mutate(se = sd / sqrt(n),
         lower = mean - qt(1 - (0.05 / 2), n - 1) * se,
         upper = mean + qt(1 - (0.05 / 2), n - 1) * se) %>%
  select(-sd, -n, -se,) %>%
  mutate(Experiment="Experiment 3", 
         Condition="Visual") %>%
  rename(distance = concept_dist)

joint_exp3 <- exp3_good %>%
  filter(condition == "joint") %>%
  group_by(concept_dist, sound_dist) %>%
  summarise(mean = mean(answer)) %>%
  mutate(Experiment="Experiment 3", 
         Condition="Joint")

exp_uni_data <- bind_rows(sounds_exp1, concepts_exp1,
                     sounds_exp2, concepts_exp2,
                     sounds_exp3, concepts_exp3)

```

Language uses symbols expressed in one modality (e.g., the auditory modality, in the case of speech) to communicate about the world, which we perceive through many different sensory modalities. Consider hearing someone yell "bee!" at a picnic, as a honey bee buzzes around the food. Identifying a word involves processing the auditory information as well as other perceptual signals (e.g., the visual image of the bee, the sound of its wings, the sensation of the bee flying by your arm). A word is successfully identified when information from these modalities provide convergent evidence. However, word identification takes place in a noisy world, and the cues received through each modality may not provide a definitive answer. On the auditory side, individual acoustic word tokens are almost always ambiguous with respect to the particular sequence of phonemes they represent, which is due to the inherent variability of how a phonetic category is realized acoustically [@hillenbrand1995]. And some tokens may be distorted additionally by mispronunciation or ambient noise. Perhaps the speaker was yelling "pea" and not "bee". Similarly, a sensory impression may not be enough to make a definitive identification of a visual category.\footnote{In the general case, language can of course be visual as well as auditory, and object identification can be done through many modalities. For simplicity, we focus on audio-visual matching here.} Perhaps the insect was a beetle or a fly instead.

How does the mind deal with uncertainty to identify the speaker's intended word? To address this question, many researchers in recent years have adopted the rational approach to cognition [@anderson90], which corresponds to Marr's computational level of analysis. This approach has had a tremendous impact not only on speech related research [@Norris08; @clayard08; @feldman2009; @kleinschmidt2015], but also on many other disciplines in the cognitive sciences [for reviews, see @chater06; @Knill04; @tenenbaum11]. Within this framework, researchers typically consider an "ideal observer" model which provides a characterization of the task/goal and shows what the optimal performance should be under this characterization. When there is uncertainty in the input, the ideal observer performs an optimal probabilistic inference. For example, in order to recognize an ambiguous linguistic input, the model uses all available probabilistic knowledge in order to maximize the accuracy of this recognition. Thus, when the task is well specified, the ideal observer model can be seen as a theoretical upper limit on performance. It is often used, not so much as a realistic model of human performance, as much as a baseline against which human performance can be compared [@Geisler2003]. When there is a deviation from the ideal, this can reveal extra constraints on human cognition, such as limitations on working memory or attentional resources.

In this line of research, @clayard08 simulated auditory uncertainty by manipulating the probability distribution of a cue (VOT) that differentiated similar words (e.g., "beach" and "peach"). They found that humans were sensitive to these probabilistic cues and their judgments closely reflected the optimal predictions. In another work, @feldman2009 studied the perceptual magnet effect, which is related to how categories influence the way the stimuli are perceived. They showed that this effect can be explained as the consequence of optimally solving the problem of perception under uncertainty. Both studies explored optimal performance under uncertainty from one modality, i.e., the auditory modality. There is however extensive evidence that information from the visual modality, in particular the speaker's facial features,  influences the perception of auditory information [see @Campbell2008 for a review]. @bejjanki2011 offered a mathematical characterization of how cues from speech and lip movements can be optimally combined and showed that human performance during audio-visual phonemic labeling was consistent (at least at the qualitative level) with the behavior of the ideal observer.

The speaker's facial features are not the only visual information that can be used to disambiguate the auditory input, however. For instance, in a referential context, information about the identity of the semantic referent can be integrated with linguistic information to resolve lexical and syntactic ambiguities in speech [e.g., @Eberhard1995; @Tanenhaus1995; @spivey2002]. To our knowledge, however, no study offered an ideal observer analysis of word identification in such context, that is, when the mind has to combine cues from the sound and the referent. Imagine, for example, that someone is uncertain whether they heard "pea" or "bee", does this uncertainty make them rely more on the referent (e.g., the object being pointed at)? Vice versa, if they are not sure if they saw a bee or a fly, does it make them rely more on the sound? More importantly, when input in both modalities is uncertain to varying degrees, do they weight each modality according to its relative reliability (which is the optimal strategy), or do they over-rely on a particular modality (which is a sub-optimal strategy)?

@@

<!-- Outside the realm of language per se, research on multisensory integration suggests that the brain integrates information from different modalities in a statistically optimal way, that is, the integration results in a higher overall precision, and modalities are weighed according to their relative reliability [@ernst02]. However, it is not a priori obvious whether this finding can be extended to word-referent recognition under multimodal uncertainty.  Indeed, there seems to be at least two fundamental differences between these two cases, and both can influence the cue combination strategy. -->

First, in the case of audio-visual speech, modalities are understood to differ only in terms of their reliability. In a referential context, however, modalities also differ in terms of their roles in the referential process: the auditory input represents the symbol (i.e., the word) whereas the visual input represents the referent. It has been suggested that because of its referential property, speech is a privileged signal for humans, starting in infancy [see @vouloumanos2014 for a review].\footnote{There is a debate over whether speech is privileged for infant and adults for the same reasons. Whereas Waxman and colleagues suggest that speech is privileged for both infants and adults because of its ability to refer, Sloutsky and colleagues suggest that speech might not have a referential status from the start. Rather, according to Sloutsky and colleagues, speech seems to be prefered by infants only because of a low level auditory "overshadowing".} Thus, in a referential context, it is possible that we do not treat the auditory and visual modalities as equivalent sources of information. Instead, there could be a bias for the auditory modality beyond what is expected from informational reliability alone.

<!--An example of such integration is when we determine the size of an object using the visual and haptic modalities [@ernst02], or when we determine the spacial location of a stimulus using the visual and the auditory modalities [@alais04] -->
Second, in the case of audio-visual speech, the associations are expected to be perceptually correlated. The expectation for this correlation is such that when there is a mismatch between auditory and visual input, people still integrate them into a unified (but illusory) percept [@mcgurk1976]. In the case of referential language, however, the multimodal association is by nature *arbitrary* [@saussure1916; @greenberg1957].  For instance, there is no logical/perceptual connection between the sound "bee" and the corresponding insect. Moreover, variation in the way the sound "bee" is pronounced is generally not expected to correlate perceptually with variation in the shape (or any other visual property) in the category of bees. Thus, one could imagine that cue combination in the case of arbitrary audio-visual associations (word-referent) is less automatic, more effortful, and therefore less conducive to optimal integration than it is in the case of perceptually correlated associations (as in audio-visual speech perception).

In the current study, we investigate how people combine cues from the auditory and the visual modality to recognize words in a referential context. In particular, we study how this combination is performed under various degrees of uncertainty in both the auditory and the visual modality.  We perform a rational analysis of the task. First we propose an ideal observer model that performs the combination in an optimal fashion. Second we compare the predictions of the optimal model to human responses. Humans can deviate from the ideal for several reasons. For instance, as mentioned above, a sub-optimality can be induced by the suggested privileged status of speech or by the arbitrariness of the referential association.  In order to study possible patterns of sub-optimality, we compare the optimal model (which provides a normative benchmark) to a descriptive model which is identical to the optimal model, except that the parameters in the former are fitted to human responses. <!--We do not perform the ideal observer analysis at the individual level, but at the population level. There might be some individual variation, but this study addresses the overall performance in the population as a whole.-->

We tested the ideal observer model's predictions in three behavioral experiments where we varied the source of uncertainty. In Experiment 1, audio-visual tokens were ambiguous with respect to their category membership only. In Experiment 2, we intervened by adding background noise to the auditory modality, and in Experiment 3, we intervened by adding background noise to the visual modality. In all experiments, participants were quantitatively near-optimal, though overall response precision was slightly lower than expected. Moreover, in Experiment 1 where neither of the modalities was perturbed with background noise, participants weighed auditory and visual cues according to the relative reliability predicted by the optimal model. In other words, we found no evidence for a modality bias towards either the auditory or the visual modality. However, in Experiment 2 and 3, participants over-relied on one modality when the other modality was perturbed with additional noise. <!-- In the discussion, we talk about the broad implications of these results and especially in relation to some developmental findings in early word learning [e.g., @stager1997]. -->

<!--The nature of the multi-modal associations (redundant vs. arbitrary) influences the quality of multi-modal processing in both children and adults: the processing and integration is easier in the case of redundant input, than it in the case of arbitrary input [see @robinson2010 for a review]. Thus, there is a priori no obvious reason multimodal cue combination in the case of arbitrary assoiations (such as the case of word-referent) should be optimal in the same way cue combination of redundant associations is. -->
<!--

A few studies have explored some aspects of audio-visual processing in a probabilistic framework  [e.g., @bejjanki2011; @kleinschmidt2015]. In these studies, the researchers focused on the specific case of audio-visual speech  where information is *redundant* across modalities. In the present work, we study the case of word reference where the audio-visual associations are usually *arbitrary* (Cite Saussure). For instance, whereas in the case of audio-visual speech, the sound "bee" and the corresponding lip movements are supposed to merge into a unified percept, there is nothing in the sound "bee" or in the way it can be pronounced, that correlates perceptually with the shape (or some other visual property) of the insect.

More generally, we know that the nature of the multi-modal associations (redundant vs. arbitrary) influences the quality of the processing in both children and adults [see @robinson2010 for a review]. In the case of redundant multimodal information (e.g. determining the size of an object through the visual and haptic modality), the sensory integration is often statistically optimal (Ernst and Banks, 2002). Optimality in this context means two things: first, the precision of the bimodal estimate increases compared to the unimodal estimates, and second, the cues from different modalities are combined according to their relative reliability. In the case of arbitrary associations, empirical studies suggest, rather, sub-optimality. In fact, processing of the bimodal input does not increase the precision, it tends to hinder performance instead (Sloutsky, Palmer, Stager, ... I should dig into the literature on the arbitrary associations). Moreover, modalities are not optimally weighted according to their reliability, instead, one modality dominates the other depending on the task, e.g., Sloutsky et al. reported auditory dominance (for kids) and then prefernce for adults (..), and Colovita et al. reported visual dominance for adults.

-->

# Models and predictions

@@

## The Audio-Visual Word Recognition Task

We introduce a new task that tests word recognition in a referential context.  We use two visual categories (cat and dog) and two auditory categories (/b/ and /d/ embedded in the minimal pair /aba/-/ada/). For each participant, an arbitrary pairing is set between the auditory and the visual categories, leading to two audio-visual word categories (e.g., dog-/aba/, cat-/ada/). In each trial, participants are presented with an audio-visual target (the prototype of the target category), immediately followed by an audio-visual test stimulus (Figure\ \@ref(fig:task)). The test stimulus may differ from the target in both the auditory and the visual components.  After these two presentations, participants press "same" or "different".

```{r task, fig.cap = "Overview of the task. In the audio-visual condition, participants are first presented with an audio-visual target (the prototype of the target category), immediately followed by an audio-visual test. The test may differ from the target in both the auditory and the visual components. After these two presentations, participants press `same' (i.e., the same category as the target) or `different' (not the same category). The auditory-only and visual-only conditions are similar to the audio-visual condition, except that only the sounds are heard, or only the pictures are shown, respectively.", fig.align = "center", out.width = "400px"}
knitr::include_graphics("pictures/task.png", dpi = 108)
```

>>>>

\documentclass[english,man]{apa6}

\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}

% Table formatting
\usepackage{longtable, booktabs}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}

\newenvironment{lltable}
  {\begin{landscape}\begin{center}\begin{ThreePartTable}}
  {\end{ThreePartTable}\end{center}\end{landscape}}

  \usepackage{ifthen} % Only add declarations when endfloat package is loaded
  \ifthenelse{\equal{\string man}{\string man}}{%
   \DeclareDelayedFloatFlavor{ThreePartTable}{table} % Make endfloat play with longtable
   % \DeclareDelayedFloatFlavor{ltable}{table} % Make endfloat play with lscape
   \DeclareDelayedFloatFlavor{lltable}{table} % Make endfloat play with lscape & longtable
  }{}%

% The following enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand\getlongtablewidth{%
 \begingroup
  \ifcsname LT@\roman{LT@tables}\endcsname
  \global\longtablewidth=0pt
  \renewcommand\LT@entry[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}%
  \@nameuse{LT@\roman{LT@tables}}%
  \fi
\endgroup}

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            pdfauthor={},
            pdftitle={Word-Referent Identification Under Multimodal Uncertainty},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=black,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setlength{\parindent}{0pt}
%\setlength{\parskip}{0pt plus 0pt minus 0pt}

\setlength{\emergencystretch}{3em}  % prevent overfull lines

\ifxetex
  \usepackage{polyglossia}
  \setmainlanguage{}
\else
  \usepackage[english]{babel}
\fi

% Manuscript styling
\captionsetup{font=singlespacing,justification=justified}
\usepackage{csquotes}
\usepackage{upgreek}

 % Line numbering
  \usepackage{lineno}
  \linenumbers

\usepackage{tikz} % Variable definition to generate author note

% fix for \tightlist problem in pandoc 1.14
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Essential manuscript parts
  \title{Word-Referent Identification Under Multimodal Uncertainty}

  \shorttitle{Word Identification Under Multimodal Uncertainty}

  \author{Abdellah Fourtassi\textsuperscript{1}~\& Michael C. Frank\textsuperscript{1}}

  % \def\affdep{{"", ""}}%
  % \def\affcity{{"", ""}}%

  \affiliation{
    \vspace{0.5cm}
          \textsuperscript{1} Department of Psychology, Stanford University  }

  \authornote{
    Abdellah Fourtassi

@@

    Correspondence concerning this article should be addressed to Abdellah
    Fourtassi, Postal address. E-mail:
    \href{mailto:afourtas@stanford.edu}{\nolinkurl{afourtas@stanford.edu}}
  }

  \abstract{Identifying a spoken word in a referential context requires both the
ability to process and integrate multimodal input and the ability to
reason under uncertainty. How do these tasks interact with one another?
We introduce a task that allows us to examine how adults identify words
under joint uncertainty in the auditory and visual modalities. We
propose an ideal observer model of the task which provides an account of
how auditory and visual cues are combined optimally. Model predictions
are tested in three experiments where word recognition is made under two
kinds of uncertainty: category ambiguity and/or distorting noise. In all
cases, the optimal model explains much of the variance in human mean
judgments. In particular, when the signal is not distorted with noise,
participants weight the auditory and visual cues optimally, that is,
according to the relative reliability of each modality. But when one
modality has noise added to it, human perceivers systematically prefer
the unperturbed modality to a greater extent than the optimal model
does. The study provides a formal framework which helps us understand
precisely how word form and word meaning interact in word recognition
under uncertainty. Moreover it offers a first step towards a model that
accounts for form-meaning synergy in early word learning.}
  \keywords{Language understanding; audio-visual processing; word learning; speech
perception; computational modeling. \\

    \indent Word count: X
  }

  \usepackage[sortcites=false,sorting=none]{biblatex}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
\newtheorem{proposition}{Proposition}
\theoremstyle{definition}
\newtheorem{example}{Example}
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}

\maketitle

\setcounter{secnumdepth}{0}

Language uses symbols expressed in one modality (e.g., the auditory
modality, in the case of speech) to communicate about the world, which
we perceive through many different sensory modalities. Consider hearing
someone yell \enquote{bee!} at a picnic, as a honey bee buzzes around
the food. Identifying a word involves processing the auditory
information as well as other perceptual signals (e.g., the visual image
of the bee, the sound of its wings, the sensation of the bee flying by
your arm). A word is successfully identified when information from these
modalities provide convergent evidence. However, word identification
takes place in a noisy world, and the cues received through each
modality may not provide a definitive answer. On the auditory side,
individual acoustic word tokens are almost always ambiguous with respect
to the particular sequence of phonemes they represent, which is due to
the inherent variability of how a phonetic category is realized
acoustically (Hillenbrand, Getty, Clark, \& Wheeler, 1995). And some
tokens may be distorted additionally by mispronunciation or ambient
noise. Perhaps the speaker was yelling \enquote{pea} and not
\enquote{bee}. Similarly, a sensory impression may not be enough to make
a definitive identification of a visual
category.\footnote{In the general case, language can of course be visual as well as auditory, and object identification can be done through many modalities. For simplicity, we focus on audio-visual matching here.}
Perhaps the insect was a beetle or a fly instead.

How does the mind deal with uncertainty to identify the speaker's
intended word? To address this question, many researchers in recent
years have adopted the rational approach to cognition (Anderson, 1990),
which corresponds to Marr's computational level of analysis. This
approach has had a tremendous impact not only on speech related research
(Clayards, Tanenhaus, Aslin, \& Jacobs, 2008; Feldman, Griffiths, \&
Morgan, 2009; Kleinschmidt \& Jaeger, 2015; Norris \& McQueen, 2008),
but also on many other disciplines in the cognitive sciences (for
reviews, see Chater \& Manning, 2006; Knill \& Pouget, 2004; Tenenbaum,
Kemp, Griffiths, \& Goodman, 2011). Within this framework, researchers
typically consider an \enquote{ideal observer} model which provides a
characterization of the task/goal and shows what the optimal performance
should be under this characterization. When there is uncertainty in the
input, the ideal observer performs an optimal probabilistic inference.
For example, in order to recognize an ambiguous linguistic input, the
model uses all available probabilistic knowledge in order to maximize
the accuracy of this recognition. Thus, when the task is well specified,
the ideal observer model can be seen as a theoretical upper limit on
performance. It is often used, not so much as a realistic model of human
performance, as much as a baseline against which human performance can
be compared (Geisler, 2003). When there is a deviation from the ideal,
this can reveal extra constraints on human cognition, such as
limitations on working memory or attentional resources.

In this line of research, Clayards et al. (2008) simulated auditory
uncertainty by manipulating the probability distribution of a cue (VOT)
that differentiated similar words (e.g., \enquote{beach} and
\enquote{peach}). They found that humans were sensitive to these
probabilistic cues and their judgments closely reflected the optimal
predictions. In another work, Feldman et al. (2009) studied the
perceptual magnet effect, which is related to how categories influence
the way the stimuli are perceived. They showed that this effect can be
explained as the consequence of optimally solving the problem of
perception under uncertainty. Both studies explored optimal performance
under uncertainty from one modality, i.e., the auditory modality. There
is however extensive evidence that information from the visual modality,
in particular the speaker's facial features, influences the perception
of auditory information (see Campbell, 2008 for a review). Bejjanki,
Clayards, Knill, and Aslin (2011) offered a mathematical
characterization of how cues from speech and lip movements can be
optimally combined and showed that human performance during audio-visual
phonemic labeling was consistent (at least at the qualitative level)
with the behavior of the ideal observer.

The speaker's facial features are not the only visual information that
can be used to disambiguate the auditory input, however. For instance,
in a referential context, information about the identity of the semantic
referent can be integrated with linguistic information to resolve
lexical and syntactic ambiguities in speech (e.g., Eberhard,
Spivey-Knowlton, Sedivy, \& Tanenhaus, 1995; Spivey, Tanenhaus,
Eberhard, \& Sedivy, 2002; Tanenhaus, Spivey-Knowlton, Eberhard, \&
Sedivy, 1995). To our knowledge, however, no study offered an ideal
observer analysis of word identification in such context, that is, when
the mind has to combine cues from the sound and the referent. Imagine,
for example, that someone is uncertain whether they heard \enquote{pea}
or \enquote{bee}, does this uncertainty make them rely more on the
referent (e.g., the object being pointed at)? Vice versa, if they are
not sure if they saw a bee or a fly, does it make them rely more on the
sound? More importantly, when input in both modalities is uncertain to
varying degrees, do they weight each modality according to its relative
reliability (which is the optimal strategy), or do they over-rely on a
particular modality (which is a sub-optimal strategy)?

@@

First, in the case of audio-visual speech, modalities are understood to
differ only in terms of their reliability. In a referential context,
however, modalities also differ in terms of their roles in the
referential process: the auditory input represents the symbol (i.e., the
word) whereas the visual input represents the referent. It has been
suggested that because of its referential property, speech is a
privileged signal for humans, starting in infancy (see Vouloumanos \&
Waxman, 2014 for a
review).\footnote{There is a debate over whether speech is privileged for infant and adults for the same reasons. Whereas Waxman and colleagues suggest that speech is privileged for both infants and adults because of its ability to refer, Sloutsky and colleagues suggest that speech might not have a referential status from the start. Rather, according to Sloutsky and colleagues, speech seems to be prefered by infants only because of a low level auditory "overshadowing".}
Thus, in a referential context, it is possible that we do not treat the
auditory and visual modalities as equivalent sources of information.
Instead, there could be a bias for the auditory modality beyond what is
expected from informational reliability alone.

Second, in the case of audio-visual speech, the associations are
expected to be perceptually correlated. The expectation for this
correlation is such that when there is a mismatch between auditory and
visual input, people still integrate them into a unified (but illusory)
percept (McGurk \& MacDonald, 1976). In the case of referential
language, however, the multimodal association is by nature
\emph{arbitrary} (Greenberg, 1957; Saussure, 1916). For instance, there
is no logical/perceptual connection between the sound \enquote{bee} and
the corresponding insect. Moreover, variation in the way the sound
\enquote{bee} is pronounced is generally not expected to correlate
perceptually with variation in the shape (or any other visual property)
in the category of bees. Thus, one could imagine that cue combination in
the case of arbitrary audio-visual associations (word-referent) is less
automatic, more effortful, and therefore less conducive to optimal
integration than it is in the case of perceptually correlated
associations (as in audio-visual speech perception).

In the current study, we investigate how people combine cues from the
auditory and the visual modality to recognize words in a referential
context. In particular, we study how this combination is performed under
various degrees of uncertainty in both the auditory and the visual
modality. We perform a rational analysis of the task. First we propose
an ideal observer model that performs the combination in an optimal
fashion. Second we compare the predictions of the optimal model to human
responses. Humans can deviate from the ideal for several reasons. For
instance, as mentioned above, a sub-optimality can be induced by the
suggested privileged status of speech or by the arbitrariness of the
referential association. In order to study possible patterns of
sub-optimality, we compare the optimal model (which provides a normative
benchmark) to a descriptive model which is identical to the optimal
model, except that the parameters in the former are fitted to human
responses.

We tested the ideal observer model's predictions in three behavioral
experiments where we varied the source of uncertainty. In Experiment 1,
audio-visual tokens were ambiguous with respect to their category
membership only. In Experiment 2, we intervened by adding background
noise to the auditory modality, and in Experiment 3, we intervened by
adding background noise to the visual modality. In all experiments,
participants were quantitatively near-optimal, though overall response
precision was slightly lower than expected. Moreover, in Experiment 1
where neither of the modalities was perturbed with background noise,
participants weighed auditory and visual cues according to the relative
reliability predicted by the optimal model. In other words, we found no
evidence for a modality bias towards either the auditory or the visual
modality. However, in Experiment 2 and 3, participants over-relied on
one modality when the other modality was perturbed with additional
noise.

\section{Models and predictions}\label{models-and-predictions}

@@

\subsection{The Audio-Visual Word Recognition
Task}\label{the-audio-visual-word-recognition-task}

We introduce a new task that tests word recognition in a referential
context. We use two visual categories (cat and dog) and two auditory
categories (/b/ and /d/ embedded in the minimal pair /aba/-/ada/). For
each participant, an arbitrary pairing is set between the auditory and
the visual categories, leading to two audio-visual word categories
(e.g., dog-/aba/, cat-/ada/). In each trial, participants are presented
with an audio-visual target (the prototype of the target category),
immediately followed by an audio-visual test stimulus
(Figure~\ref{fig:task}). The test stimulus may differ from the target in
both the auditory and the visual components. After these two
presentations, participants press \enquote{same} or \enquote{different}.

\begin{figure}

{\centering \includegraphics[width=400px]{pictures/task}

\caption{Overview of the task. In the audio-visual condition, participants are first presented with an audio-visual target (the prototype of the target category), immediately followed by an audio-visual test. The test may differ from the target in both the auditory and the visual components. After these two presentations, participants press `same' (i.e., the same category as the target) or `different' (not the same category). The auditory-only and visual-only conditions are similar to the audio-visual condition, except that only the sounds are heard, or only the pictures are shown, respectively.}\label{fig:task}
\end{figure}
